\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-05-14 doc/sw-system.tex
abstract+concl. and few other edits in between

Strong lenses are expensive to find, because they are rare. The highest purity
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{BoltonEtal2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{NegrelloEtal2010}. Such searches have to yield pure samples,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields of only tens to hundreds of lenses. An alternative
approach is to search images already of sufficiently high resolution and colour
contrast, and confirm the systems as gravitational lenses by modeling the
survey data themselves \citep[][]{MarshallEtal2009}. Several square degrees of
HST images have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{MoustakasEtal2007,FaureEtal2008,Jackson2008,MoreEtal2011,
PawaseEtal2014}. Similarly, searches of over a hundred square degrees of CFHT
Legacy Survey (\cfhtls) ground-based imaging, also with sub-arcsecond image quality,
have revealed a smaller number of wider image separation group-scale systems
\citep[e.g.][]{CabanacEtal2007,MoreEtal2012}. Detecting galaxy-scale lenses
from the ground is difficult, but feasible albeit with lower efficiency and requiring
HST or spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{GavazziEtal2014}.

@@

The four CFHT Legacy Survey\footnote{\texttt{http://www.cfht.hawaii.edu/Science/CFHTLS/}}
\citep[\cfhtls,][]{Gwyn2012} ``Wide'' fields cover a total of approximately
160 square degrees of sky (after taking into account tile overlaps).
With high and homogeneous image quality (the mean seeing in the $g$-band is
0.78''), and reaching limiting magnitudes of around 25 across the $ugriz$ filter
set, this survey has yielded several dozen new gravitational lenses on both
galaxy and group scales
\citep{GavazziEtal2014,SonnenfeldEtal2013a,CabanacEtal2007,MoreEtal2012}. The
quality of the data, combined with the presence of these comparison ``known
lens'' samples, makes this a natural choice against which to develop and test
the \SW system. The \cfhtls is also well-representative of the data quality
expected from several next-generation sky surveys, such as DES, KiDS, HSC and
LSST. We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}} for this
work.

@@

At \StageTwo, where no retirement was carried out, it was possible to reach
100\% purity: the knee of the curve is at just under 90\% completeness. However,
the purity decreases rapidly if higher completeness than this is sought. The
optimal sample in this simulated lens search experiment would have been
constructed with a threshold value of $\pr(\LENS|C,T) > 0.47$. At 100\% purity
and 90\% completeness, it would have contained around 89 lens candidates.
However, we remind the reader that all these values are dependent on the
properties of the training subjects, which were chosen to be fairly
visible: we expect the completeness to real lenses to be somewhat lower, as a
result.

@@

How fast does the \SW crowd classify subjects? Each software agent records the timestamp
of each classification its volunteer makes; by measuring the time lags between
successive classifications, we can make estimates of the crowd's classification
and contribution speed. We plot the former quantity for both \StageOne and
\StageTwo in \Fref{fig:results:speed}, normalizing the speed and time axes to
their respective totals. The fractional classification rates
in each stage of the survey fall off in approximately the same way, despite the
factor of nearly 20 difference in crowd size. \StageOne (consisting of
$\sim430,000$ subjects) was completed in around 5000 hours (by some 37,000
participants), with classification rates in the first few days averaging at
$\sim10^4$ per hour. This was stimulated by various forms of advertising (press
releases, and emails to registered Zooniverse users). The asymptotic
classification rate was  around ten times lower, at $\sim10^3$ per hour.
Interestingly, the average skill was found to be approximately constant over the
lifetime of \StageOne, leading to a contribution rate that tracks closely the
classification rate. (A 10\% increase in average skill was seen over the first
40 days, and a decrease of the same amount in the last 40 days -- not enough to
cause a significant difference between classification and contribution rate
behavior, but perhaps reflecting a learning period at the start and then a
decrease
in participation late on.) \StageTwo (3700 subjects) was completed in around 600
hours by around 2000 participants, with classification rates starting at
$~\sim2000$ per hour and decaying to $\sim100$ per hour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=\linewidth]{rate_classification.pdf}
\caption{Fractional classification rate in the \SW \cfhtls survey. Fractional classification rate is the number
of classifications per unit time, divided by the final total number of
classifications. Fractional survey time is the time elapsed since the beginning
of the survey, divided by the total length of the survey. The unit of time in
the fractional  classification rate is one survey, so that the curves in
integrals under the curves are both one. }
\label{fig:results:speed}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We investigated the source of the incompleteness and impurity visible in
\Fref{fig:results:sample:CP}, by examining the \StageTwo false negatives and
false positives, and their behaviour as they are classified using the online
analysis trajectory plots introduced in \Sref{sec:swap}
(\Fref{fig:swap:subject-trajectories}).
\Fref{fig:discuss:performance:trajectories} shows 2 example simulated lenses
that were missed ($\pr(\LENS|C,T) < 10^{-5}$) by the \SW system (top row),
and 2 example
non-lenses that were incorrectly flagged as candidates ($\pr(\LENS|C,T) >
0.95$) by the \SW system (bottom row).

@@

We have designed, implemented and tested a system for   detecting new strong
gravitational lens candidates in wide field imaging surveys by crowd-sourced
visual inspection. The \SW web-based classification interface presents carefully
prepared colour composite sky images to volunteers, who mark features they
propose to  have been lensed. The participants receive ongoing training with a
mixture of simulated lens and known-to-be-empty images, and we use this
information to inform the automated interpretation of their markers. In our
first lens search we simply divided the \cfhtls imaging into some 430,000 tiles,
and collected over 11~million volunteer image classifications.

@@

\item Participation levels were high (over 36,000 volunteers, contributing
classifications at rates between $10^3$ and $10^4$ images per hour), suggesting
that if this can be maintained, visual inspection of tens of thousands of images
could be performed just a few weeks. An expanded crowd of $10^6$ volunteers
(which has already been reached in the Galaxy Zoo project) would
be able to inspect a plausible sample of $10^6$ LSST or Euclid targets
on similar time scales.

@@

\item The ``skill'' of the volunteers correlates strongly with the number of
images they have classified.  Quantifying the ``contribution'' as the integrated
skill over the test subjects classified, we find that about 90\% of the
contribution was made by a few hundred volunteers ($1\%$ of the crowd) -- but
that the width of the skill distribution means that a  few thousand volunteers
could have made comparable contributions had they classified more images.

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91-93\%$ and $<1\%$ (Stages 1 and
2). This FPR translates to an approximate purity of 30\%. We find a
higher purity at \StageTwo (perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).

\item The simulated gravitational lenses that were missed were predominantly
galaxy-scale lenses with faint blue galaxy sources, whose lensed features are
difficult to distinguish from the light from the lens galaxy. We observed some
additional scatter, with some simulated lenses accumulating low probability
simply due to classification noise.

>>>>

Strong lenses are expensive to find, because they are rare. The highest purity
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{BoltonEtal2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{NegrelloEtal2010}. Such searches have to yield pure samples,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields of only tens to hundreds of lenses. An alternative
approach is to search images already of sufficiently high resolution and colour
contrast, and confirm the systems as gravitational lenses by modeling the
survey data themselves \citep[][]{MarshallEtal2009}. Several square degrees of
HST images have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{MoustakasEtal2007,Jackson2008,MoreEtal2011,
PawaseEtal2014}. Similarly, searches of over a hundred square degrees of CFHT
Legacy Survey (\cfhtls) ground-based imaging, also with sub-arcsecond image quality,
have revealed a smaller number of wider image separation group-scale systems
\citep[e.g.][]{CabanacEtal2007,MoreEtal2012}. Detecting galaxy-scale lenses
from the ground is difficult, but feasible albeit with lower efficiency and requiring
HST or spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{GavazziEtal2014}.

@@

The four CFHT Legacy Survey\footnote{\texttt{http://www.cfht.hawaii.edu/Science/CFHTLS/}}
\citep[\cfhtls,][]{Gwyn2012} ``Wide'' fields cover a total of approximately
160 square degrees of sky (after taking into account tile overlaps).
With high and homogeneous image quality (the mean seeing in the $g$-band is
0.78''), and reaching limiting magnitudes of around 25 across the $ugriz$ filter
set, this survey has yielded several dozen new gravitational lenses on both
galaxy and group scales
\citep{GavazziEtal2014,SonnenfeldEtal2013a,CabanacEtal2007,MoreEtal2012}. The
quality of the data, combined with the presence of these comparison ``known
lens'' samples, makes this a natural choice against which to develop and test
the \SW system. The \cfhtls is also well-representative of the data quality
expected from several next-generation sky surveys, such as DES, KiDS, HSC and
LSST. We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{\texttt{ftp://t07.terapix.fr/pub/T07/}} for this
work.

@@

At \StageTwo, where no retirement was carried out, it was possible to reach
100\% purity: the knee of the curve is at just under 90\% completeness. However,
the purity decreases rapidly if higher completeness than this is sought. The
optimal sample in this simulated lens search experiment would have been
constructed with a threshold value of $\pr(\LENS|C,T) > 0.47$. At 100\% purity
and 90\% completeness, it would have contained around 89 lens candidates.
However, we remind the reader that all these values are dependent on the
properties of the training subjects, which were chosen to be fairly
visible: we expect the completeness to real lenses to be somewhat lower, as a
result (see \PaperTwo).

@@

How fast does the \SW crowd classify subjects? Each software agent records the timestamp
of each classification its volunteer makes; by measuring the time lags between
successive classifications, we can make estimates of the crowd's classification
and contribution speed. We plot the former quantity for both \StageOne and
\StageTwo in \Fref{fig:results:speed}, normalizing the speed and time axes to
their respective totals. The fractional classification rates
in each stage of the survey fall off in approximately the same way, despite the
factor of nearly 20 difference in crowd size. \StageOne (consisting of
$\sim430,000$ subjects) was completed in around 5000 hours (by some 37,000
participants), with classification rates in the first few days averaging at
$\sim10^4$ per hour. This was stimulated by various forms of advertising (press
releases, and emails to registered Zooniverse users). The asymptotic
classification rate was  around ten times lower, at $\sim10^3$ per hour.
Interestingly, the average skill was found to be approximately constant over the
lifetime of \StageOne, leading to a contribution rate that tracks closely the
classification rate. (A 10\% increase in average skill was seen over the first
40 days, and a decrease of the same amount in the last 40 days -- not enough to
cause a significant difference between classification and contribution rate
behavior, but perhaps reflecting a learning period at the start and then a
decrease
in participation later on.) \StageTwo (3400 subjects) was completed in around 600
hours by around 2000 participants, with classification rates starting at
$~\sim2000$ per hour and decaying to $\sim100$ per hour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=\linewidth]{rate_classification.pdf}
\caption{Fractional classification rate in the \SW$-$\cfhtls survey. Fractional classification rate is the number
of classifications per unit time, divided by the final total number of
classifications. Fractional survey time is the time elapsed since the beginning
of the survey, divided by the total length of the survey. The unit of time in
the fractional  classification rate is one survey, so that the curves in
integrals under the curves are both one. }
\label{fig:results:speed}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We investigated the source of the incompleteness and impurity visible in
\Fref{fig:results:sample:CP}, by examining the \StageTwo false negatives and
false positives, and their behaviour as they are classified using the online
analysis trajectory plots introduced in \Sref{sec:swap}
(\Fref{fig:swap:subject-trajectories}).
\Fref{fig:discuss:performance:trajectories} shows 2 example simulated lenses
that were missed ($\pr(\LENS|C,T) < 10^{-7}$) by the \SW system (top row),
and 2 example
non-lenses that were incorrectly flagged as candidates ($\pr(\LENS|C,T) >
0.95$) by the \SW system (bottom row).

@@

We have designed, implemented and tested a system for   detecting new strong
gravitational lens candidates in wide field imaging surveys by crowd-sourced
visual inspection. The \SW web-based classification interface presents carefully
prepared colour composite sky images to volunteers, who mark features they
propose to  have been lensed. The participants receive ongoing training with a
mixture of simulated lens and known-to-be-empty images, and we use this
information to automate the interpretation of their markers. In our
first lens search we simply divided the \cfhtls imaging into some 430,000 tiles,
and collected over 11~million volunteer image classifications.

@@

\item Participation levels were high (about 37,000 volunteers, contributing
classifications at rates between $10^3$ and $10^4$ images per hour), suggesting
that if this can be maintained, visual inspection of tens of thousands of images
could be performed in just a few weeks. An expanded crowd of $10^6$ volunteers
(which has already been reached in Zooniverse) would
be able to inspect a plausible sample of $10^6$ LSST or Euclid targets
on similar time scales.

@@

\item The ``skill'' of the volunteers correlates strongly with the number of
images they have classified.  Quantifying the ``contribution'' as the integrated
skill over the test subjects classified, we find that about 90\% of the
contribution was made by a few hundred volunteers ($1\%$ of the crowd) -- but
that the broad width of the skill distribution means that a few thousand volunteers
could have made comparable contributions had they classified more images.

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91-93\%$ and $<1\%$
(Stages I and II). This FPR translates to an approximate purity of 30\%. We find a
higher purity at \StageTwo (perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).

\item The simulated gravitational lenses that were missed were predominantly
galaxy-scale lenses with faint blue galaxy sources, whose lensed features are
difficult to distinguish from the light from the lens galaxy (consistent
with what we find also for real lenses, see \PaperTwo). We observed some
additional scatter, with some simulated lenses accumulating low probability
simply due to classification noise.
